### 2013-02-19 article/paper.tex
more on mysql implementation and redis features

In this case, the algorithms were implemented as a SQL stored-procedures only,
using indices and intermediate tables when possible. A Python script was used
to run these stored-procedures and collect the benchmarks. For a storage-engine 
we chose InnoDB, the default engine in MySQL 5.5, as it out-performs MyISM in 
most senarios. \footnote{Sara: can we cite Oracle white papers? 
http://www.oracle.com/partners/en/knowledge-zone/mysql-5-5-innodb-myisam-522945.pdf}
\linebreak
Implementation details:
\begin{itemize}
	\item {\bf Helper Tables} We generated the following helper tables 
		to speed up some algorithms:
		\begin{itemize}
			\item {common-neighbors (id1, id2, neighbor)}
			\item {common-neighbors-count (id1, id2, number of common neighbors)}
			\item {neighbors (id, number of neighbors)}
			\item {topN (id, number of neighbors)}
				This table is simply the top 101 items of the neighbors table.
		\end{itemize}
	\item {\bf Common Neighbors (global)} 
		To find the top N records with most common neighbors, 
		we selected them from the common-neighbors-count table.
	\item {\bf Jaccard's Coefficient (global)} 
		We used the following relation between the cardinality of union and intersection:
		\[{|N(x) \cup N(y)|} = {|N(x)| + |N(y)| - |N(x) \cap N(y)|}\]
		So to calculate the score we used the common-neighbors-count and the neighbors helper tables.
	\item {\bf Adamic/Adar (global)} 
		We used the common-neighbors joined with the neighbors helper table.
	\item {\bf Preferential attachment (Global)}
		We used only the topN helper table, because the highest products 
		must be combinations of the nodes with most neighbors.
	\item {\bf Common Neighbors (for node)}
		We used a simple join on the edges table, no helper tables required.
	\item {\bf Jaccard's Coefficient (for node)}
		We used the relation between the cardinality of union and intersection,
		with join of two edges tables, and two neighbors helper tables.
	\item {\bf Adamic/Adar (for node)}
		We used join on the edges table, with the neighbors helper table.
	\item {\bf Preferential attachment (for node)}
		We used the topN helper table, just like in the global version.
	\item {\bf Graph Distance (for node)}
		We used a recursive procedure, starting with a temporary table with all the node's neighbors,
		and on each iteration N we added all the neighbors of nodes in depth N-1. This implementation
		required us to reconfigure MySQL maximum recursion depth.
	\item {\bf Katz (for node)}
		We used a recursive procedure to calculate $|P^l(x,y)|$ for all reachable nodes in depth 3,
		and stored that in a temporary table. We then used that to calculate the score.
	\item {\bf Rooted PageRank (for node)}

\end{itemize}

\subsection{Redis}
Redis is an open-source, in-memory key-value store database system, sponsored
by VMware. It's main features are atomic operations, data structure as value
(hash, list, set and sorted-set), optional persistence, pub/sub mechanism,
built-in Lua interpreter and master-slave replication.

Here, some of the algorithms were implemented in Python using redis API, and
some were implemented as Lua scripts (to minimize Redis API calls). When an
index was needed, we simply used another Redis DB (you get 16 with the default
configuration). 
The graphs were represented using adjacency lists, i.e. a key is a node-id and
it's value is a set of his neighbors. That implementation posed a limitation,
since we couldn't store nodes without neighbors, as redis doesn't allow to
store empty sets. \#TODO: why doesn't it matter in our case

>>>>

In this case, the algorithms were implemented as a SQL stored-procedures only,
using indices and intermediate tables when possible. A Python script was used
to run these stored-procedures and collect the benchmarks. For a storage-engine 
we chose InnoDB, the default engine in MySQL 5.5, as it out-performs MyISM in 
most senarios. \footnote{Sara: can we cite Oracle white papers? 
http://www.oracle.com/partners/en/knowledge-zone/mysql-5-5-innodb-myisam-522945.pdf}
\linebreak
Implementation details:
\begin{itemize}
	\item {\bf Helper Tables} 
        We generated the following helper tables to speed up some algorithms:
		\begin{itemize}
			\item {common-neighbors (id1, id2, neighbor)}
			\item {common-neighbors-count (id1, id2, number of common neighbors)}
			\item {neighbors (id, number of neighbors)}
			\item {topN (id, number of neighbors)}
				This table is simply the top 101 items of the neighbors table.
		\end{itemize}
	\item {\bf Common Neighbors (global)} 
		To find the top N records with most common neighbors, 
		we selected them from the common-neighbors-count table.
	\item {\bf Jaccard's Coefficient (global)} 
		We used the following relation between the cardinality of union and intersection:
		\[{|N(x) \cup N(y)|} = {|N(x)| + |N(y)| - |N(x) \cap N(y)|}\]
		So to calculate the score we used the common-neighbors-count and the neighbors helper tables.
	\item {\bf Adamic/Adar (global)} 
		We used the common-neighbors joined with the neighbors helper table.
	\item {\bf Preferential attachment (Global)}
		We used only the topN helper table, because the highest products 
		must be combinations of the nodes with most neighbors.
	\item {\bf Common Neighbors (for node)}
		We used a simple join on the edges table, no helper tables required.
	\item {\bf Jaccard's Coefficient (for node)}
		We used the relation between the cardinality of union and intersection,
		with join of two edges tables, and two neighbors helper tables.
	\item {\bf Adamic/Adar (for node)}
		We used join on the edges table, with the neighbors helper table.
	\item {\bf Preferential attachment (for node)}
		We used the topN helper table, just like in the global version.
	\item {\bf Graph Distance (for node)}
		We used a recursive procedure, starting with a temporary table with all the node's neighbors,
		and on each iteration N we added all the neighbors of nodes in depth N-1. This implementation
		required us to reconfigure MySQL maximum recursion depth.
	\item {\bf Katz (for node)}
		We used a recursive procedure to calculate $|P^l(x,y)|$ for all reachable nodes in depth 3,
		and stored that in a temporary table. We then used that to calculate the score.
	\item {\bf Rooted PageRank (for node)}
        We used an iterative loop over a temporary table <id, rpr, new-rpr>, initiated with all the 
        nodes in the graph, and 1/N value where N is the number of nodes in the graph. On every 
        iteration we updated the new-rpr value, using the formula above. We stopped when the algorithm
        converged on the top-10 scoring nodes and their order.
\end{itemize}

\subsection{Redis}
Redis is an open-source, in-memory key-value store database system, sponsored
by VMware. It's written in C and has atomic operations, we used the following features:
\begin{itemize}
	\item {\bf Transactions}
        You can use redis transaction to save time spent on network trafic 
        between the client API and the server. We used that feature to speed up
        the graph load significantly.
	\item {\bf Data structure as value}
        Redis support storing many types of data structures as valuse and 
        performe efficient, atomic operation over them. I.e. there are strings,
        hashes (key-value pairs), lists, sets (in which we used to store the 
        adjacency lists), and sorted-sets (like sets, but with rank for each item).
        Operations include: incrementing value in hash, appending to lists, 
        union/intersection/difference on sets, and getting sorted-set member with top rank.
	\item {\bf Lua Scripting}
        Lua scripting enable you to run your own code, in an atomic matter inside the server.
        Unlike transactions, you can add non-Redis logic to this atomic operation. We used 
        Lua scripting to speed-up and algorithem which took long time to run.\footnote{Sara: we are kind of abusing this feature, by locking the db and assuming single client, should we mention that?}
\end{itemize}

Here, some of the algorithms were implemented in Python using redis API, and
some were implemented as Lua scripts (to minimize Redis API calls). When an
index was needed, we simply used another Redis DB (you get 16 with the default
configuration). 
The graphs were represented using adjacency lists, i.e. a key is a node-id and
it's value is a set of his neighbors. That implementation posed a limitation,
since we couldn't store nodes without neighbors, as redis doesn't allow to
store empty sets. This isn't such an issue for us, as all the algorithems we tested
return an empty set for such nodes, and we still kept a nodes db <id, name> which
could provide us with the node count (i.e. needed for the Rooted PageRank implementation)
\linebreak
Implementation details:
\begin{itemize}
	\item {\bf Helper Database} 
        We only generated one database speed up the Preferential attachment algorithm:
	    topN (id, number of neighbors), a database which holds 101 nodes with most neighbors.
	\item {\bf Common Neighbors (global)} 
        We implemented this algorithm using the Lua scripting feature. TODO
	\item {\bf Jaccard's Coefficient (global)} 
	\item {\bf Adamic/Adar (global)} 
	\item {\bf Preferential attachment (Global)}
	\item {\bf Common Neighbors (for node)}
	\item {\bf Jaccard's Coefficient (for node)}
	\item {\bf Adamic/Adar (for node)}
	\item {\bf Preferential attachment (for node)}
	\item {\bf Graph Distance (for node)}
	\item {\bf Katz (for node)}
	\item {\bf Rooted PageRank (for node)}
\end{itemize}
